### 2020-05-27 writeup/transfer.tex

\subsection{De-noising by Transferring from a Source Task}\label{sec_similarity}

A folklore intuition since the seminal work of Caruana \cite{C97} is that task relatedness determines how well multi-task learning performs.
We formalize the intuition using our setup and technical tools.
For a simple example of two tasks, we show that whether $\hat{\beta}_t^{\MTL} \le \hat{\beta}_t^{\STL}$ is determined by $\norm{\beta_1 - \beta_2}^2$, the distance of the task models.
We derive a sharp threshold when positive transfer transitions to negative transfer, as a function of model distance $\norm{\beta_1 - \beta_2}^2$, data size $n_1$ abd $n_2$ and the noise level $\sigma$.

@@

\textbf{Implications.}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
We derive a sharp transition from positive to negative transfer as a parameter of $\sigma_1^2$.

\subsection{Improving Labeled Data Efficiency}\label{sec_data_size}

In classical Rademacher or VC based theory of multi-task learning, adding more labeled data improves the generalization performance of a model.
On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using Example \ref{ex_basic}, we analyze the effect of varying the source task data size.

@@

\textbf{Implications.} We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized by a metric called the data efficiency ratio as follows.
Given several tasks, let $\alpha^{\star}$ be the largest factors such that the total number of labeled datapoints needed for solving all the tasks can be reduced by an $\alpha^{\star}$ factor (compared to training independently) while keeping the performance nearly the same.
More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
If we only use $\alpha n_i$ datapoints from every task to train the multi-task learning estimator $\hat{\beta}(\alpha)$, then $\alpha \in (0, 1)$ will be the smallest number such that
\[ \alpha^{\star} \define\argmin_{\alpha\in(0, 1)} ~~ \te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))\le \te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL}). \]
We quantify the data efficiency ratio of $\hat{\beta}_t^{\MTL}$ for Example \ref{ex_basic} as follows.

\subsection{When to Align Task Covariances under Covariate Shift}\label{sec_covariate}

So far we have considered settings where $\Sigma_1 = \Sigma_2$.
This setting is relevant for multi-class image classification settings, where different tasks share the same input features.
In general, the covariance matrices of the two tasks may be different, e.g. in text classification.
In this part, we use our tools to provide a case study on the effect of applying multi-task learning for two tasks when $\Sigma_1 \neq \Sigma_2$.

@@

\textbf{Implications.}

%\subsection{Extension to Transfer Learning}\label{sec_taskonomy}

>>>>

\subsection{Benefits of Multi-task Learning}

\textbf{De-noising by Transferring from a Source Task}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
We derive a sharp transition from positive to negative transfer as a parameter of $\sigma_1^2$.

\textbf{Improving Labeled Data Efficiency}
We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized by a metric called the data efficiency ratio as follows.
Given several tasks, let $\alpha^{\star}$ be the largest factors such that the total number of labeled datapoints needed for solving all the tasks can be reduced by an $\alpha^{\star}$ factor (compared to training independently) while keeping the performance nearly the same.
More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
If we only use $\alpha n_i$ datapoints from every task to train the multi-task learning estimator $\hat{\beta}(\alpha)$, then $\alpha \in (0, 1)$ will be the smallest number such that
\[ \alpha^{\star} \define\argmin_{\alpha\in(0, 1)} ~~ \te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))\le \te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL}). \]
We quantify the data efficiency ratio of $\hat{\beta}_t^{\MTL}$ for Example \ref{ex_basic} as follows.

\subsection{Contributing causes of negative transfer}

A folklore intuition since the seminal work of Caruana \cite{C97} is that task relatedness determines how well multi-task learning performs.
We formalize the intuition using our setup and technical tools.

\textbf{Task model dissimilarity.}
For a simple example of two tasks, we show that whether $\hat{\beta}_t^{\MTL} \le \hat{\beta}_t^{\STL}$ is determined by $\norm{\beta_1 - \beta_2}^2$, the distance of the task models.
We derive a sharp threshold when positive transfer transitions to negative transfer, as a function of model distance $\norm{\beta_1 - \beta_2}^2$, data size $n_1$ abd $n_2$ and the noise level $\sigma$.

@@

\textbf{Data size.}
In classical Rademacher or VC based theory of multi-task learning, adding more labeled data improves the generalization performance of a model.
On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.
Using Example \ref{ex_basic}, we analyze the effect of varying the source task data size.

@@

\textbf{Covariate shift.}
So far we have considered settings where $\Sigma_1 = \Sigma_2$.
This setting is relevant for multi-class image classification settings, where different tasks share the same input features.
In general, the covariance matrices of the two tasks may be different, e.g. in text classification.
In this part, we use our tools to provide a case study on the effect of applying multi-task learning for two tasks when $\Sigma_1 \neq \Sigma_2$.

@@
